---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(kableExtra)
```

# DATASET 2 - Principal Component Analysis

## Introduction 

The weekly rates of return for five stocks listed on the New York Stock Exchange are analyzed in this report. The Stock-price data consists of 103 weekly rates of return on 5 stocks. The weekly rates of return for five stocks (JP Morgan, Citibank, Wells Fargo, Royal Dutch Shell, and ExxonMobil) listed on the New York Stock Exchange were determined for the period January 2004 through December 2005. The weekly rates of return are defined as (current week closing price-previous week closing price)/(previous week closing price), adjusted for stock splits and dividends. The source of the dataset is from Applied Multivariate Statistical Analysis, 5th Edition as T08_04.DAT


## Summary

We notice ther is linear correlation anonges all stocks. Citi bank, JP Morgan and Wells Fargo stocks are highly correlated. 

```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
stock <-  read.table("T8-4.DAT")

colnames(stock) <- c("JP Morgan","Citibank", "Wells Fargo","Royal Dutch Shell","Exxon")

ggpairs(stock, lower = list(continuous = wrap("smooth", alpha = 0.4, size = 0.3), discrete = "blank", combo="blank"), diag = list(discrete="barDiag", continuous = wrap("densityDiag", alpha=0.5 )), upper = list(combo = wrap("box_no_facet", alpha=0.5), continuous = wrap("cor", size=4, alignPercent=0.8))) + theme(panel.grid.major = element_blank())


```

## Analysis

For the principal component analysis, we start with computing the covariance matrix.

```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
# Sample covariance matrix
scov <- cov(stock)
 scov_frame <- data.frame(scov)
kable(scov_frame, format = "pandoc", caption = "Covariance matrix Forbes Dataset")
```
```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
#Principal components

stock.pca <- princomp(stock, cor=TRUE)
stock.pce <- data.frame(matrix(rbind(c(1.561, 1.18, 0.707, 0.632, 0.505 ), c(0.4874546, 0.2814025, 0.1001025, 0.08000632, 0.05103398), c(0.4874546, 0.7688572, 0.8689597, 0.94896602, 1.00000000)), nrow = 3, ncol = 5))
colnames(stock.pce) = c("comp 1", "comp 2", "comp 3","comp 4","comp 5")
rownames(stock.pce) = c("Standard deviation", "Proportion of Variance", "Cumulative Proportion ")
kable(stock.pce, digits = 4, format = "pandoc", caption = "Table 1: Principal components") 
```

We comput the scree plot to determin the number of principal components to use. From the plot, we could see that we could use the first 3 components.

```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
#Scree plot to determine number of PC to use

# A scree plot:
plot(1:(length(stock.pca$sdev)),  (stock.pca$sdev)^2, type='b', 
     main="Scree Plot", xlab="Number of Components", ylab="Eigenvalue Size")
day = factor(rep(c("M", "Tu","W","Th","F"),21))
day = day[1:dim(stock)[1]]

```
```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
plot(stock.pca$scores[,1], stock.pca$scores[,2], 
     xlab="PC 1", ylab="PC 2",  lwd=2, col=day)
legend("topright",legend=levels(day),pch=1,col=1:3,cex=0.7)
```

We could see from the plot above that there is no cluster for any particular day.

```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
biplot(stock.pca,xlabs=day)
```

Looking at figure, we could see that JP Morgan, Wells fargo and Citibank are clustered together towrds Mondays, Tuesdays and Wednesdays while Royal Dutch and Exxon are clustered together in a separate region of Friday.

```{r echo=FALSE, message=FALSE, warning= FALSE, result=FALSE}
biplot(stock.pca, choices=3:4,xlabs=day)

```


We dont see much in the plot as the third and fourth principal components do not explain a lot of the variation.

## Conclusion

We performed principal component analysis of stock data. We found that the first three principal components capture most of the variance, in fact 87%.